BACI of biomass change for VOCC source cells for all years

### 2019-09-26 analysis/VOCC/4-vocc-BACI-multiyear.Rmd
BACI matching, if do threshold must be >0.4, save plots of matches

#### Set age class
```{r echo=TRUE}
age <- "mature" # or mostly so
# age <- "imm"
```

#### Set vector type
```{r echo=TRUE}
biomass_threshold <- "50" # filter for climate values predicted to support 50% of max biomass 
 biomass_threshold <- "25" # filter for climate values predicted to support 50% of max biomass 
global_only <- TRUE

 climate <- "temperature"
 temp_threshold <- c(0.25)

# temp_threshold <- c(0.5)
# temp_threshold <- c(0.75)
#  climate <- "do"
# #  do_threshold <- c(0.5)
#  do_threshold <- c(0.25)
# climate <- "do and temperature"
```

@@

params <- params[params$species == species, ]

if (global_only) {
 params <- params %>% mutate(global = ifelse(grepl("_", model), "F", "T"))
 params <- filter(params, global == "T")
} else {
 #params <- filter(params, ssid %in% model_ssid)
}

@@

  upper_thresholds <- round(.upper[1], digits = 2)
  lower_thresholds <- NA

  variable_names <- c("temp")

@@

if (climate == "do") {
  # optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
  # optimal_dos <- na.omit(optimal_dos)
  # optimal_do <- round(optimal_dos[1], digits = 2)
  # optimal_values <- optimal_do
  params <- filter(params, lower_do_25 > 0.5)

@@

  .lower <- na.omit(.lower)
  lower_thresholds <- round(.lower[1], digits = 2)
  upper_thresholds <- NA
  variable_names <- c("do_est")

@@

  variable_names <- c("do_est", "temp")
  lower_thresholds <- c(lower_thresholds, NA)
  upper_thresholds <- c(NA, upper_thresholds)
  lower_change <- c(do_threshold, Inf)
  upper_change <- c(Inf, temp_threshold)
  indices <- c(1, 2)
  delta_t_total <- 2
  }

trim_thresholds <- c(lower_thresholds, upper_thresholds)
min_string <- paste0(lower_change, collapse = "-")
max_string <- paste0(upper_change, collapse = "-")

@@

### Filter vectors to match optimal range for chosen species
```{r}

trimmed_vectors <- lapply(vocc_by_year, function(vocc) {    
#start_time <- vocc$start_time[1]

@@

```{r}
png(
  file = paste0("figs/", spp,
    "/vocc-", min_string, "to", max_string, "-", 
    climate_string, "-", spp, "-", age, "-", ssid_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs.png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(
  grobs = trimmed_vector_plots,
  ncol = 2,
  top = grid::textGrob(paste("Vectors for cells exceeding", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass"), gp=grid::gpar(fontsize=16))
)
dev.off()
```

@@

```{r}
rm(vocc_by_sp)
try({
  vocc_by_sp <- readRDS(paste0(
    "data/ BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-multiyear-all.rds"
  ))
})

@@

match_formula <- paste0("treat ~ trawled + x + y") 
  # depth.st + omega_s + muddy.st + any_rock.st + 
  # for (j in (variable_names)) {
  #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
  #   }

@@

  # need to be standardized? seems not
  alldata$depth2 <- alldata$depth^2
  alldata$start_density.st <- round(arm::rescale(log(alldata$start_density))*5)/5
  alldata$depth.st <- round(arm::rescale(log(alldata$depth))*5)/5
  alldata$muddy.st <- arm::rescale(alldata$muddy) 
  alldata$any_rock.st <- arm::rescale(alldata$any_rock) 
  #alldata$omega_s.st <- arm::rescale(alldata$omega_s)

@@

  matched <- MatchIt::matchit(as.formula(match_formula), 
    exact = c("start_density.st", "start_time", "depth.st"),
    method = "nearest", m.order = "random",
    distance = "probit",
    discard = "both",
    data = alldata)
  # matched
# })
}
```

@@

#  data
#})
vocc_by_sp <- list()
vocc_by_sp[[1]] <- data
vocc_by_sp[[3]] <- match_formula  
vocc_by_sp[[4]] <- match_plots
saveRDS(vocc_by_sp, file = paste0(
    "data/", spp,
    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-multiyear-all.rds"
  ))
}
```

## Save spp dataframe
```{r}
#biomass_threshold <- "50"
optimal_string <- paste0(na.omit(trim_thresholds), collapse = "-")
#climate <- "temperature"
try({
  vocc_by_sp <- readRDS(paste0(
    "data/", spp,
    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-multiyear-all.rds"
  ))
})

n <- nrow(vocc_by_sp[[1]])
species <- rep(species, n) 
data <- vocc_by_sp[[1]]
df <- cbind(species, data)
species <- species[[1]]

@@

# View(data)
```

@@

ggplot(data, 
#ggplot(filter(data, as.numeric(matchobs) <= 550), 
  aes(y=Y, x=X, shape=as.factor(treat), colour=as.factor(matchobs))) +
  geom_point(alpha=0.8, size = 1) +
  scale_shape_manual(values= c(0, 15)) +
  scale_colour_viridis_d(option="C", guide = FALSE) +
  facet_wrap(~start_time) +
  theme_bw()

@@

ggplot(data, aes(std_log_depth, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~start_time) +
  theme_bw()
#  
# ggplot(data, aes(y=log_density, x=depth, colour=cell_type)) +
#   geom_point(alpha=0.35) +
#   scale_colour_viridis_d(option="C") +
#     facet_grid(start_time~time) +
#   theme_bw()

ggplot(data, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_grid(start_time~time, scales = 'fixed') +
  theme_bw()

ggplot(data, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha=0.45) +
  scale_fill_viridis_d(option="C") +
  facet_grid(time~start_time) +
  theme_bw()

@@

ggplot(data, aes(y=log_density, x=cell_type, fill=cell_type)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  geom_text(data = dat_summary,
            aes(cell_type, max(data$log_density)-1, label = n), hjust = -0.3) +
  facet_grid(time~start_time) +
  theme_bw()
```

```{r}
# png(
#   file = paste0("figs/", spp,
#     "/matches-", min_string, "to", max_string, "-", 
#     climate_string, "-", spp, "-", age, "-", ssid_string, "-", 
#     climate_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs.png"
#   ),
#   res = 600,
#   units = "in",
#   width = 8.5,
#   height = 11
# )
# gridExtra::grid.arrange(
#   grobs = ,
#   ncol = 2,
#   top = grid::textGrob(paste("Matched pairs for", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass"), gp=grid::gpar(fontsize=16))
# )
# dev.off()
```
```

### Basic BACI
```{r}
for (j in (variable_names)) {
  baci_formula <- paste0("log_density~after + cell_type + after:cell_type + n_targets + vect_dist + vect_dist2 + std_", j, "+ std_", j, "2 + (1|icell) + (1|matchobs) + (1|start_time)")
}

@@

```{r}
sjPlot::plot_model(m, type = "std2", order.terms=c(1,2,8,3,4,5,6,7), title = paste(species, "adult"))
```

### GAMM
```{r}
 gamm_formula <- paste0("log_density ~ ", 
  # "time * start_time + ",
  #"start_time * cell_type + ",
  "time * cell_type + s(vect_dist) + s(std_log_depth) + s(X,Y)")

@@

gam4 <- gamm4::gamm4(as.formula(gamm_formula), 
  random = ~(1|matchobs) + (1|icell) 
   + (1|start_time)
  # + (time|start_time) 
  # + (cell_type|start_time)
  , data=data)

@@

>>>>

#### Set age class
```{r echo=TRUE}
age <- "mature" # or mostly so
age <- "imm" 
```

#### Set vector type
```{r echo=TRUE}
global_only <- TRUE
biomass_threshold <- "75"

# biomass_threshold <- "50" # filter for climate values predicted to support 50% of max biomass 
#  biomass_threshold <- "25" # filter for climate values predicted to support 25% of max biomass... doesn't work for DO

climate <- "temperature"
 temp_threshold <- c(0.25)
# temp_threshold <- c(0.5)
# temp_threshold <- c(0.75)
# temp_threshold <- c(1)

 # climate <- "do"
 # do_threshold <- c(0.25)
# do_threshold <- c(0.5)

# climate <- "do and temperature"
```

@@

params <- params[params$species == species, ] 
glimpse(params)

if (global_only) {
 params <- params %>% mutate(global = ifelse(grepl("_", model), "F", "T"))
 params <- filter(params, global == "T")
} else {
 if (model_ssid == 4) {
   params <- params %>% mutate(survey4 = ifelse(grepl("4", model), "T", "F"))
   params <- filter(params, survey4 == "T")
 }
}

@@

  upper_thresholds <- round(.upper[1], digits = 2)
  upper_model <- params[rev(order(params$t_maxy)), ]$model

  lower_thresholds <- NA
  lower_model <- NULL
  variable_names <- c("temp")

@@

if (climate == "do") {
  # optimal_dos <- params[rev(order(params$do_maxy)), ]$optimal_do
  # optimal_dos <- na.omit(optimal_dos)
  # optimal_do <- round(optimal_dos[1], digits = 2)
  # optimal_values <- optimal_do
  params <- filter(params, lower_do_25 > 0.4)

@@

  .lower <- na.omit(.lower)
  lower_thresholds <- round(.lower[1], digits = 2)
  lower_model <- params[rev(order(params$do_maxy)), ]$model
  upper_thresholds <- NA
  upper_model <- NULL
  variable_names <- c("do_est")

@@

  variable_names <- c("do_est", "temp")
  lower_thresholds <- c(lower_thresholds, NA)
  upper_thresholds <- c(NA, upper_thresholds)
  lower_change <- c(do_threshold, Inf)
  upper_change <- c(Inf, temp_threshold)
  upper_model <- params[rev(order(params$t_maxy)), ]$model
  lower_model <- params[rev(order(params$do_maxy)), ]$model

  indices <- c(1, 2)
  delta_t_total <- 2
  }

trim_thresholds <- c(lower_thresholds, upper_thresholds)
trim_models <- c(lower_model, upper_model)

min_string <- paste0(lower_change, collapse = "-")
max_string <- paste0(upper_change, collapse = "-")

@@

### Filter vectors to match optimal range for chosen species
```{r}
trimmed_vectors <- lapply(vocc_by_year, function(vocc) {    
#start_time <- vocc$start_time[1]

@@

```{r}
png(
  file = paste0("figs/", spp,
    "/vocc-", min_string, "to", max_string, "-", 
    climate_string, "-", spp, "-", age, "-", ssid_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs.png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(
  grobs = trimmed_vector_plots,
  ncol = 2,
  top = grid::textGrob(paste("Vectors for cells exceeding", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass (", trim_models, ")"), gp=grid::gpar(fontsize=16))
)
dev.off()
```

@@

```{r}
rm(vocc_by_sp)
# try({
#   vocc_by_sp <- readRDS(paste0(
#     "data/BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
#   ))
# })

@@

match_formula <- paste0("treat ~ trawled + x + y + depth_st") 
  # depth.st + omega_s + muddy.st + any_rock.st + 
  for (j in (variable_names)) {
    match_formula <- paste0("", match_formula, " + ", j, "_s.st")
    start_value.st <- paste0("", j, "_s.st")
    start_value <- paste0("", j, "_s")

    alldata <- alldata %>% mutate((!!start_value.st) := 
        arm::rescale(alldata[[start_value]]) )
  #(UQ(rlang::sym(paste0(variable_names[j], "_s")))))
    }

@@

  # need to be standardized? seems not
  alldata$depth2 <- alldata$depth^2
  alldata$start_density.st <- round(arm::rescale(log(alldata$start_density))*5)/5
  alldata$depth_st <- arm::rescale(log(alldata$depth))
  alldata$depth.st <- round(arm::rescale(log(alldata$depth))*5)/5
  alldata$muddy.st <- arm::rescale(alldata$muddy) 
  alldata$any_rock.st <- arm::rescale(alldata$any_rock)

@@

  matched <- MatchIt::matchit(as.formula(match_formula), 
    exact = c("start_density.st", "start_time"),
    method = "nearest", m.order = "random",
    distance = "probit",
    discard = "both",
    data = alldata)
  # matched
# })
}
```

@@

#  data
#})
vocc_by_sp <- list()
vocc_by_sp[[1]] <- data
vocc_by_sp[[3]] <- match_formula  
vocc_by_sp[[4]] <- match_plots
saveRDS(vocc_by_sp, file = paste0(
    "data/", spp,
    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
  ))
}
```

## Save spp dataframe
```{r}
#biomass_threshold <- "50"
optimal_string <- paste0(na.omit(trim_thresholds), collapse = "-")
#climate <- "temperature"
try({
  vocc_by_sp <- readRDS(paste0(
    "data/", spp,
    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
  ))
})

n <- nrow(vocc_by_sp[[1]])
species <- rep(species, n) 
mature <- rep(age, n)
data <- vocc_by_sp[[1]]
class <- cbind(species, mature)
df <- cbind(class, data)
species <- species[[1]]

@@

# View(df)
```

@@

p1 <- ggplot(data, 
#ggplot(filter(data, as.numeric(matchobs) <= 550), 
  aes(y=Y, x=X, shape=as.factor(treat), colour=as.factor(matchobs))) +
  geom_point(alpha=0.8, size = 1) +
  scale_shape_manual(values= c(0, 15)) +
  scale_colour_viridis_d(option="C", guide = FALSE) +
  facet_wrap(~start_time) +
  theme_bw()
p1

@@

p2 <- ggplot(data, aes(std_log_depth, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~start_time) +
  theme_bw()
p2

# ggplot(data, aes(y=log_density, x=depth, colour=cell_type)) +
#   geom_point(alpha=0.35) +
#   scale_colour_viridis_d(option="C") +
#     facet_grid(start_time~time) +
#   theme_bw()

p3 <- ggplot(data, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_grid(start_time~time, scales = 'fixed') +
  theme_bw() + theme(legend.position='none')
p3

p4 <- ggplot(data, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha=0.45) +
  scale_fill_viridis_d(option="C") +
  facet_grid(time~start_time) +
  theme_bw() + theme(legend.position='none')
p4

@@

p5 <- ggplot(data, aes(y=log_density, x=cell_type, fill=cell_type)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  geom_text(data = dat_summary,
            aes(cell_type, min(data$log_density)+1, label = n), hjust = 1.25) +
  facet_grid(time~start_time) +
  theme_bw() + theme(legend.position='none')
p5
```

```{r}
png(
  file = paste0("figs/", spp,
    "/matches-", min_string, "to", max_string, "-",
    climate_string, "-", spp, "-", age, "-", ssid_string, "-",
    climate_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs1.png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(
  p1,p2,p3,p4,p5,
  ncol = 2,
  top = grid::textGrob(paste("Matched pairs for", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass"), gp=grid::gpar(fontsize=16))
)
dev.off()
```

### Basic BACI
```{r eval=F}
for (j in (variable_names)) {
  baci_formula <- paste0("log_density~after + cell_type + after:cell_type + n_targets + vect_dist + vect_dist2 + std_", j, "+ std_", j, "2 + (1|icell) + (1|matchobs) + (1|start_time)")
}

@@

```{r eval=F}
sjPlot::plot_model(m, type = "std2", order.terms=c(1,2,8,3,4,5,6,7), title = paste(species, "adult"))
```

### GAMM
```{r eval=F}
 gamm_formula <- paste0("log_density ~ ", 
  # "time * start_time + ",
  #"start_time * cell_type + ",
  "time * cell_type  + s(std_log_depth) + s(X,Y) ") #+ s(vect_dist)

@@

gam4 <- gamm4::gamm4(as.formula(gamm_formula), 
  random = ~(1|matchobs) + (1|icell) 
  # + (1|start_time)
  + (time|start_time) 
  # + (cell_type|start_time)
  , data=data)

@@

```{r}
spp
```
